The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-01-18 paper/paper.tex
Paper text

% Add your own macros here:
\pdfoutput=1 %for arXiv submission
%\usepackage{amsmath,amssymb,amstext}
\usepackage{amsmath,amstext}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
\usepackage[figure,figure*]{hypcap}

\usepackage{url}
\urlstyle{same}

@@

\begin{itemize}
    \item Goal of challenge: identify binning method(s) leading to best cosmology constraints 
\end{itemize}

@@

Metric 2 uses a figure of merit based on that presented by the Dark Energy Task Force \citep{detf}.  We 
use the inverse area of a Fisher Matrix ellipse, which approximates the size in two axes 
of the posterior PDF one would obtain on analysing the spectra:
\begin{align}
    F &= \left( \frac{\partial \mu}{\partial \theta} \right)^T C^{-1} \left( \frac{\partial \mu}{\partial \theta} \right), \\
    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{\Omega_c,\sigma_8})}}}
\end{align}
where $[F^{-1}]_{\Omega_c,\sigma_8}$ extracts the 2x2 matrix for these two parameters.

@@

\subsection{Control Methods}
\begin{itemize}
    \item single bin
    \item one galaxy per bin?
    \item sanity check: random selection should be same as single bin
\end{itemize}

\section{Methods}
\begin{itemize}
    \item Gradient Boosting \@ jecampagne
    \item Neural Network \@EiffL
    \item Random Forest
    \item \placeholder{cuts in photometry space without redshift estimation \@barber?}
\end{itemize}

\section{Results}
\begin{itemize}
    \item Big table of results
    \item fisher matrix ellipses overplotted for top N methods
    \item Plots of n(z)
\end{itemize}

\section{Discussion}
\begin{itemize}
    \item Overall effectiveness of method classes
    \item Individual winner
    \item Largest number of bins that proves workable
\end{itemize}

@@

@@

\begin{itemize}
    \item Infrastructure for checking submissions
    \item Guidance for users regarding caching and batch processing
    \item Timing
    \item Training set size
    \item Advertisement
\end{itemize}

>>>>

% Add your own macros here:
\pdfoutput=1 %for arXiv submission
%\usepackage{amsmath,amssymb,amstext}
\usepackage{amsmath,amstext}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
\usepackage[figure,figure*]{hypcap}
\usepackage{natbib}
\usepackage{url}
\urlstyle{same}

@@

@@

Metric 2 uses a figure of merit based on that presented by the Dark Energy Task Force \citep{detf}.  We 
use the inverse area of a Fisher Matrix ellipse, which approximates the size in two axes 
of the posterior PDF one would obtain on analysing the spectra:
\begin{align}
    F &= \left( \frac{\partial \mu}{\partial \theta} \right)^T C^{-1} \left( \frac{\partial \mu}{\partial \theta} \right), \\
    S &= \frac{1}{2 \pi \sqrt{\det{([F^{-1}]_{p_1, p_2})}}}
\end{align}
where $[F^{-1}]_{p_1, p_2}$ extracts the 2x2 matrix for two parameters.  We use both the original
DETF variant in which $p_1 = w_0, p_2 = w_a$, representing the constraining power on dark energy 
evolution, and a version in which $p_1 = \Omega_c$ and $p_2 = \sigma_8$, representing constraining
power on overall cosmic structure amplitude and its evolution.

@@

\subsection{Control Methods}

Two control methods were used to test the challenge machinery and metrics.
The first method randomly assigned galaxies to each tomographic bin, resulting
in bins which each had the same $n(z)$, on average.  Using this method, we expect
that increasing the number of bins should not increase any metric, since bins are
maximally correlated.  We find this to be true for all our metrics.

Second, we use a method which uses only the $i$-band to select tomographic bins,
splitting galaxies in equal-sized bins in this magnitude. We expect only a small
increase in metrics with the number of bins, and also that the addition of the $g$-band
should make no difference to metrics.  Again, we find this to be true.

One additional method was submitted to the challenge with the caveat that it could
never be applied to real data and should therefore considered a form of control method.
This method was {\sc Utopia}, which found, for each object in the testing data, the
nearest neighbour object in the training sample.  The high density, completeness, and
representativeness of our training set made this possible; realistics spectroscopic training
sets are too sparse.  The method performed well, as expected, 
but was not the highest scoring overall.

\section{Methods}

\subsection{ {\sc RandomForest} }
This method was submitted by the challenge organizers and used a random
forest algorithm to assign galaxies.

Random forests \cite{breiman2001} train a collection of individual \emph{decision trees},
each of which consists of a set of bifurcating comparisons in which different features in
the data (in this case, magnitudes and errors) are compared to criterion values to choose
which branch to follow.  Each ``leaf'' (end comparison) selects a classification bin for
an input galaxy, and the tree is trained to choose comparison features and comparison values
which maximally split the discrimination at each step, and in final leaf purity.

Since decision trees tend to over-fit data, random forests generate a suite of decision
trees, each randomly choosing a subset of features on which to train at each bifurcation.
An average over the trained trees is taken as the overall classification.

In this implementation we chose nominal bin edges by splittng ensuring equal numbers of training set 
objects were in each bin, and then trained the forest to map magnitudes, magnitude errors, and
galaxy sizes to bin indices.  We use the {\sc scikit-learn} implementation of the random forest
\citep{scikit-learn}.

% Please describe your methods with some very general / high level background theory
% and a focus on the spsecific implementation.  Add any citations you like with bibtex in
% paper.bib
% Feel free to rename your method, but please leave a comment explaining so that I can do the
% same in the tables.

\subsection{ {\sc Autokeras LSTM} }

\subsection{ {\sc CNN} }

\subsection{ {\sc ComplexSOM} }

\subsection{ {\sc ENSEMBLE} }

\subsection{ {\sc GPzBinning} }

\subsection{ {\sc JaxCNN} }

\subsection{ {\sc JaxResNet} }

\subsection{ {\sc LGBM} }

\subsection{ {\sc LSTM} }

\subsection{ {\sc NeuralNetwork 1} }

\subsection{ {\sc NeuralNetwork 2} }

\subsection{ {\sc PCACluster} }

\subsection{ {\sc SimpleSOM} }

\subsection{ {\sc TCN} }

\subsection{ {\sc UTOPIA} }

\subsection{ {\sc ZotBin} }

\subsection{ {\sc ZotNet} }

\subsection{ {\sc funbins} }

\subsection{ {\sc mlpqna} }

\subsection{ {\sc myCombinedClassifiers} }

\section{Results}

\begin{table*}[]
\begin{tabular}{l|llll|llll|}
                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
{\sc Autokeras\_LSTM } & 31.1 & 74.9    & 103.4    & nan    & 44.1             & 67.5             & 122.2             & 98.4\\
{\sc CNN } & 27.1 & 50.8    & 76.7    & 103.3    & 30.5             & 57.7             & 95.4             & 122.4\\
{\sc ComplexSOM } & 38.0 & 52.1    & 94.4    & 101.6    & 34.9             & 45.0             & 91.6             & 100.3\\
{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
{\sc funbins } & 36.8 & 82.5    & 122.1    & 141.8    & 42.1             & 100.4             & 142.5             & 167.2\\
{\sc GPzBinning } & 26.2 & 49.7    & 80.8    & 111.7    & 27.8             & 55.3             & 87.2             & 126.9\\
{\sc IBandOnly } & 38.0 & 50.0    & 54.4    & 57.3    & 38.0             & 50.0             & 54.4             & 57.3\\
{\sc JaxCNN } & 59.9 & 101.9    & 105.2    & nan    & 79.7             & 125.2             & 150.0             & nan\\
{\sc JaxResNet } & 73.7 & 111.1    & 131.6    & nan    & 82.2             & 126.0             & nan             & 161.5\\
{\sc LGBM } & 27.0 & 50.8    & 78.8    & 108.2    & 30.3             & 57.4             & 92.9             & 125.2\\
{\sc LSTM } & 26.7 & 51.1    & 81.4    & 103.6    & 30.2             & 57.1             & 95.9             & 126.9\\
{\sc mlpqna } & 39.3 & 64.7    & 93.2    & 121.8    & 42.8             & 72.2             & 109.0             & 133.7\\
{\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
{\sc NeuralNetwork\_1 } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
{\sc NeuralNetwork\_2 } & 0.0 & 0.0    & 0.0    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
{\sc PCACluster } & 29.5 & 68.1    & 75.8    & 0.0    & 50.1             & 72.7             & 90.8             & 0.0\\
{\sc Random } & 1.2 & 1.2    & 1.2    & 1.2    & 1.2             & 1.2             & 1.2             & 1.2\\
{\sc RandomForest } & 39.5 & 65.2    & 93.2    & 118.5    & 43.0             & 72.8             & 110.1             & 136.5\\
{\sc SimpleSOM } & 39.5 & 59.9    & 77.9    & 98.5    & 42.2             & 73.2             & 108.0             & 134.8\\
{\sc TCN } & 40.0 & 65.3    & 90.9    & 110.5    & 44.1             & 72.8             & 108.0             & 134.0\\
{\sc UTOPIA } & 39.0 & 63.7    & 90.8    & 112.2    & 43.1             & 74.0             & 112.8             & 139.3\\
{\sc ZotBin } & 64.8 & 106.4    & 121.8    & 135.3    & 77.6             & 120.0             & 141.8             & 154.7\\
{\sc ZotNet } & 73.6 & 111.2    & 131.8    & 145.9    & 83.7             & 128.5             & 150.1             & 167.2\\
\end{tabular}
\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the 3x2pt DETF figure-of-merit achieved by entrant methods on the 
CosmoDC2 part of the challenge.}
\end{table*}

\begin{itemize}
    \item Big table of results
    \item fisher matrix ellipses overplotted for top N methods
    \item Plots of n(z)
\end{itemize}

\section{Discussion}
\begin{itemize}
    \item Overall effectiveness of method classes
    \item Individual winner
\end{itemize}

\subsection{Individual winners}

RUN WINNERS WITH UGRIZY TO GET AN IDEA OF THE LOSS FROM DROPPING U AND Y.

\subsection{Large bin numbers}

It is clear from values above that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.
Once bins become this narrow the primary analysis concern will become the calibration of overall
bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

The inclusion of the $g$ band typically increases metrics by around 10--20\%, including for the
highest scoring methods.  This value should be weighed against the challenge of high-accuracy
PSF measurement for this band, but until the end of the survey this is unlikely to be
the limiting factor in overall precision.

@@

\bibliography{paper}

@@

\begin{itemize}
    \item Splitting classification from nominal bin choice
    \item Infrastructure for checking submissions
    \item Guidance for users regarding caching and batch processing
    \item Timing
    \item Training set size
    \item Advertisement
\end{itemize}
